Energy and exergy analysis of a cruise ship

### 2017-10-10 Manuscript/Manuscript.tex
(almost) final changes to the code
Most of the bugs have been eliminated

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
%% Tthe amsmath package provides the "align" environment for nicer equation formatting
\usepackage{amsmath}
\def\bibsection{\section*{References}}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

@@

According to the third IMO GHG Study, in 2012 CO$_2$ emissions from shipping amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}. Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}. As an example, global transport demand has increased by 3.4\% in 2014, compared to a global GBP growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}.

International Energy Agency data show that the OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3]. The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4]. Finally, in the Baltic Sea an emission control area is enforced by the International Maritime Organisation since January 2015 which stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels. More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}

@@

Given the existing limitations of the current available information in scientific literature highlighted in the previous section, the objective of this work is the following:
\begin{itemize}
	\item Analyse the demand of a cruise ship in terms of propulsion power, electric power and heat, based on operational measurements.
	\item Analyse the current efficiency of the system, and potential ways to improve it, by means of applying exergy analysis
	\item Provide a more detailed analysis of the heat demand of a cruise ship, as a basis for further studies on how to optimize the whole efficiency of the propulsion plant.
\end{itemize}

@@

Here an overview of the method is provided: 
\begin{itemize}
	\item Data gathering
	\item Data filtering
	\item Data pre-processing
	\item Heat demand estimation
	\item Statistical analysis
	\item Energy and exergy analysis
\end{itemize}

In this paper, we present the application of energy and exergy analysis (described in sections \ref{set:met:energy} and \ref{sec:met:exergy} to a cruise ship. This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; these are used to generate the required information for calculating energy and exergy flows (Section \ref{sec:met:processing}. In these regards, the estimation of the heat demand is treated in a separate section (Sec. \ref{sec:met:heat} as it constituted a particularly challenging task and it is one of the central aspects of this article.

\subsection{Energy analysis}

@@

\subsection{Exergy analysis}

@@

\subsection{System performance}

@@

The ship systems are summarized in Figure \ref{fig:shipSystems}. The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller. The main engines are four Wärtsilä 4-stroke Diesel engines (ME) rated 5850 kW each All engines are equipped with selective catalytic reactors (SCR) for NOX emissions abatement. Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.

@@

Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska Högarna for the ambient temperature. The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska Högarna lighthouse is along the ship route, in between the Swedish archipelago and Åland. The assumption was validated based on  June-December period, for which onboard measurements were available. This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work. The fit the SMHI-data with the rest of the data-set the SMHI data was resampled from 1h for the seawater and 3h for the ambient temperature, to 15min frequency using a linear interpolation.

\subsection{Data processing: Ship energy system modeling} \label{sec:met:modeling}

@@

\subsubsection*{Main engines}

Make a figure representing the main engines, with specified the variables for which we have measurements of.

List the main assumptions

\subsubsection*{Auxiliary engines}

Make a figure representing the main engines, with specified the variables for which we have measurements of.

List the main assumptions

\subsubsection*{Auxiliary boilers}

\subsubsection*{Cooling systems}

\subsubsection*{Production-side}

In a top-down model, the demand of an energy system is modeled based on the production side of the energy balance. According to the top-down approach, the total heat consumption is calculated based on the sum of the heat provided by the three main heat sources available on board:
\begin{itemize}
	\item Exhaust gas boilers
	\item Auxiliary oil-fired boilers
	\item Heat recovery on the HT cooling systems
\end{itemize}

For the case of the EGBs, the heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
\begin{equation}
\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
\end{equation}\label{eq:egb}

Where measurements of the exhaust gas temperature before and after the EGBs (respectively $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$) are available from the DLS; the mass flow of exhaust gas ($\dot{m}_{eg}$) is calculated as described in appendices \ref{app:me} and \ref{app:ae}; and the specific heat at constant pressure of the exhaust gas ($c_{p,eg}$) is assumed to be constant and equal to 1.08 $\frac{kJ}{kgK}$.

For the case of the ABs, the heat transferred to the steam ($\dot{Q}_{AB}$) is calculated according to equation \ref{eq:ab}:
\begin{equation}
\dot{Q}_{AB} = \dot{m}_{fuel,AB} \eta_{AB}
\end{equation} \label{eq:ab}

Where the fuel consumption of the ABs ($\dot{m}_{fuel,AB}$) is available on a monthly basis. No information was available concerning the first-law efficiency of the boilers ($ \eta_{AB}$). \cite{Cohen1962} estimated it to vary between 0.83 and 0.89 depending on the load of the boiler, while according to a more recent experimental campaign presented in \cite{Mrzljak2017} the efficiency varies between 0.7 and 0.79. In the absence of more specific available data, a value of 0.8 was selected for $ \eta_{AB}$

The contribution from the heat recovery on the high temperature cooling systems of the engines ($\dot{Q}_{HRHT}$) represents the main uncertainty related to the top-down estimations. This contribution cannot be measured, neither directly or indirectly, and hence needs to be estimated.

For this reason, two calculated values for the top-down heat demand are provided:
\begin{itemize}
	\item A \textbf{high boundary} case, where it is assumed that almost all of the heat available from the HT cooling systems (an utilization factor of 0.9 is assumed to account for losses) is used. This corresponds to the assumption that the use of the waste heat from the HT cooling systems is prioritized over the use of waste heat from the exhaust gas of the engines.
	\begin{equation}
	\dot{Q}_{HRHT} = f_{HRHT} \dot{Q}_{HT,tot} 
	\end{equation}
	\item a \textbf{low boundary} case, where it is assumed that the waste heat from the cooling systems is only used when the ship is in port (again assuming a utilization factor of 0.9). 
	\item a reference case, (NOTE: We should discuss what the assumption could be here. The one we had in the ECOS paper? Or a value halfway between the highest and lowest estimate?)
\end{itemize}

In both cases, the available heat from the HT cooling systems is estimated as explained in Section \ref{eq:met:modeling}.

\subsubsection*{Consumers-side}
Following, the assumptions that we have made related to the bottom-up estimations of ship heat demand:
\begin{itemize}
	\item In the HVAC unit, the "\textbf{Preheater}" is only used when heating (i.e., in winter) while the "\textbf{Reheater}" is only used when cooling (i.e., in summer)
	\item \textbf{Spacial heating} demand during \textbf{winter} is linearly dependent on the external air temperature, from a minimum of 0 for an external temperature of 20$^o$C to a maximum of 3500 kW (design power of the preheater)  for an outer air temperature of -20$^o$C.
	\item \textbf{Spacial heating} demand during \textbf{summer} is related to the need for dehumidification, and harder to estimate based on available measurements. As a first assumption, it is estimated to be equal to the cooling demand (estimated as detailed in Sec. ref{sec:met:electric})
	\item Every passenger contributes with a total of 150 W of free heat during the day (8-23, ref. to "walking, seated" activity from \cite{Wang2000}) and 100 W during night (23-8, ref. to "seated at rest" activity from \cite{Wang2000}). For the crew, as it is assumed that they perform a more intense activity than the passengers, the daily contribution is considered equal to 160 (ref. to "moderate work" activity from \cite{Wang2000}) 
	\item The heat required for \textbf{tank heating} is linearly dependent on the external sea water temperature. The maximum heat is required for an external seawater temperature of 0$^o$C
	\item The heat required from the \textbf{galley} depends on two parameters:
	\begin{itemize}
		\item The \textbf{number of passengers} (linear dependence)
		\item the \textbf{hour of the day}. See Figure \ref{fig:hwhvstime}
	\end{itemize}
	\item The heat required from the \textbf{hot water} depends on two parameters:
	\begin{itemize}
		\item The \textbf{number of passengers} (linear dependence)
		\item the \textbf{hour of the day}. The hour-dependent factor is assumed based on the estimations for a land-based hotel reported in \cite{Chung2015}.
	\end{itemize}   
\end{itemize}

In Table \ref{tab:heatConsumers} we summarize all heat consumers on board the case study vessel. For each entry, we provide the design heat demand and the variables we assumed the actual heat demand to depend on:
\begin{table}
	\begin{tabular}{lrccccc}
		\hline 
		Name & $\dot{Q}_{des}$ [kW] & \multicolumn{5}{c}{Depends on...} \\ 
		\hline 
		 & & $T_{air}$ & t & $N_{pax}$ & $f_{TL}$ & $\lambda$ \\ 
		\hline 
		Hot water heater & 1200 &  & x & x &  &  \\ 
		HVAC Preheater & 3500 & x & x & x &  &  \\ 
		HVAC Reheater & 1780 & x & x & x &  &  \\ 
		Tank heating & 208 &  &  &  & x &  \\ 
		Other tanks & 138 &  &  &  & x &  \\ 
		HFO Tank heating & 271 &  &  &  & x &  \\ 
		Machinery space heaters & 281 & x &  &  &  &  \\ 
		Bilge water separator & 26 &  &  &  &  &  \\ 
		Hot water calorifier & 366 &  & x &  &  &  \\ 
		Fuel oil heater & 103 &  &  &  &  & x \\ 
		HFO Separator & 37 &  &  &  &  &  \\ 
		Galley & 602 &  & x & x &  &  \\
		\hline 
	\end{tabular} \label{tab:heatConsumers}
\end{table}

@@

\subsubsection*{HVAC systems}

@@

\subsubsection*{Thrusters}

When entering port areas, the ship needs to use thrusters to maneuver and berth. In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant.

The observation of Figure XX shows how the power demand from thrusters can be identified as "spikes" in the total electric power demand. In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average. From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).

@@

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
%% Tthe amsmath package provides the "align" environment for nicer equation formatting
\usepackage{amsmath}
%% The package tabls helps in having nicer tables with higher spaces
\usepackage{tabls}
%% The package multirow allows merging together multiple cells vertifcally
\usepackage{multirow}
%% 
\usepackage{subcaption}

\def\bibsection{\section*{References}}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

@@

CO$_2$ emissions from shipping in 2012 amounted to a total of 949 million tonnes, contributing to 2.7\% of global anthropogenic CO$_2$ emissions \cite{Smith2014}. Although this contribution appears relatively low, the trend is that shipping will play an even greater role in the future due to the increased transport demand according to all IMO future scenarios \cite{Smith2014}. As an example, global transport demand has increased by 3.4\% in 2014, compared to a global GBP growth of 2.5\% the same year, which shows how shipping tends to rise even faster than global economy \cite{UNCTAD2015}. The OECD countries have reduced the CO$_2$ impact from shipping, but a larger amount has been moved to the non-OECD countries [3]. The fact that shipping needs to even further reduce its CO$_2$ emissions in the near future is essential for being able to achieve the goals of maintaining the climate below a 2-degree level in 2050 [4].

In addition to considerations related to GHG emissions, an emission control area (ECA) is enforced in the Baltic Sea  by the International Maritime Organisation since January 2015. This ECA stipulates that the fuel used must not contain more than 0.1\% sulphur, therefore requiring the use of more expensive distillate fuels. More generally making shipping sustainable is a challenge that will demand growing attention by the shipping industry \cite{Andersson2016}

@@

Given the existing limitations of the current available information in scientific literature highlighted in the previous section, the objective of this work is the following:
\begin{itemize}
	\item Analyse the demand of a cruise ship in terms of propulsion power, electric power and heat, based on operational measurements. The lack of existing information with reference to the heat demand is considered a particular element of novelty when compared to existing literature in the subject.
	\item Analyse the current efficiency of the system, and potential ways to improve it, by means of applying exergy analysis.
	\item Provide reference operational conditions for further use in the field of energy systems optimization.
\end{itemize}

@@

In this paper, we present the application of energy and exergy analysis (described in section \ref{sec:met:energyExergy}) to a cruise ship. This is done for one specific case study vessel, that is described in detail in Section \ref{sec:met:case}; in Section \ref{sec:met:gathering} we present the specifics of the available information, in particular the measured data and the technical documentation; the specific assumptions and methods used for processing the available information into  energy and exergy flows are specified in Section \ref{sec:met:processing}. In these regards, the estimation of the heat demand is treated in a separate section (Sec. \ref{sec:met:heat}) as it constitutes a particularly challenging task and it is one of the central aspects of this article.

\subsection{Energy and exergy analysis} \label{sec:met:energyExergy}

\subsubsection{Energy analysis}

@@

\subsubsection{Exergy analysis}

@@

\subsubsection{System performance}

@@

The ship systems are summarized in Figure \ref{fig:shipSystems}. The propulsion system is composed of two equal propulsion lines, each made of two engines, a gearbox, and a propeller. The main engines are four Wärtsilä 4-stroke Diesel engines (ME) rated 5850 kW each. All engines are equipped with selective catalytic reactors (SCR) for NOX emissions abatement. Propulsion power is needed whenever the ship is sailing; however, it should be noted that the ship rarely sails at full speed, and most of the time it only needs one or two engines operated simultaneously.

@@

Since measurements of ambient and seawater temperature from onboard logging systems were not available for the whole dataset, we used measurements taken from SMHI database for Landsort lighthouse for the seawater, and the lighthouse Svenska H\"{o}garna for the ambient temperature. The Landsord lighthouse is situated south of the Stockholm archipelago, and the Svenska H\"{o}garna lighthouse is along the ship route, in between the Swedish archipelago and Åland. The assumption was validated based on  June-December period, for which onboard measurements were available. This resulted in a root mean square error of 1.5 K and 1.9 K for the seawater and the ambient temperature respectively, which we considered to be accurate enough for the purpose of this work. The fit the SMHI-data with the rest of the data-set the SMHI data was resampled from 1h for the seawater and 3h for the ambient temperature, to 15min frequency using a linear interpolation.

\subsection{Data processing: Ship energy system modeling} \label{sec:met:processing}

@@

\subsubsection{Diesel engines}

The setup of the engines and their connections to the other ship systems are summarized in Figure REF and REF for the main and auxiliary engines, respectively. The defining net energy flows are listed in Table REF. It should be noted that, in the case of the main engines, the engine power output was not measured and needed to be estimated. In this work, we calculated the engine power output based on measurements of engine fuel rack position (used as a proxy of the mass of fuel injected per cycle) and of the engine speed:
\begin{eqnarray}
\dot{m}_{fuel} & = & \dot{m}_{fuel,des} \left(a_0 + a_1 \frac{frp}{frp_{des}}\right) \frac{\omega}{\omega_{des}} \\
\eta_{ME} & = & a_0 + a_1 \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} + a_2 \left( \frac{\dot{m}_{fuel}}{\dot{m}_{fuel,des}} \right)^2 \\
\dot{W}_{ME} & = & \dot{m}_{fuel} \eta_{ME} LHV
\end{eqnarray}

The validity of this assumption can be seen by observing Figure REF, representing the calculated total power from the main engines as a function of the ship speed

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/DieselEngines}
	\caption{Schematic representation of the Diesel engines and their connections to the cooling systems}
	\label{fig:DieselEngines}
\end{figure}

In the case of the auxiliary engines, instead, direct measurements of the power generated by each engine were available.

For both auxiliary and main engines, part of the relevant internal variables are not measured (particularly with relation to the bypass valves) and this required to determine them based on modeling the heat and mass balance of the engine. The model is explained in detail in Appendix REF.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/Pme_vs_vship_Neng}
	\caption{Scatter plot, Propulsion power versus ship speed}
	\label{fig:Pme_vs_vship_Neng}
\end{figure}

ALSO INCLUDE THE POWER VS TC SPEED RELATIONSHIP

\begin{table}
	\small
	\centering
	\begin{tabular}{llccrrrrr}
		\hline
		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
		\hline
		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
		&   $\beta_{comp}$		& - 	& M & & & & &  \\
		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
		&   $\beta_{exp}$		& - 	& C & & & & &  \\
		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
		&	$T_{eg,out}$  		& $^o$C & M  & & & & &  \\ 
		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
		\multicolumn{2}{l}{LOC} & & & & & & &  \\
		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
		\multicolumn{2}{l}{JWC} & & & & & & &  \\
		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
		\hline
		\end{tabular}
	\caption{Main engines, measured and calculated temperatures and flows at different engine loads}
\end{table}

INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA

\begin{table}
	\small
	\centering
	\begin{tabular}{llccrrrrr}
		\hline
		\multicolumn{2}{l}{Variable} & Unit & M/C & 20\% & 40\% & 60\% & 80\% & 100\% \\
		\hline
		\multicolumn{2}{l}{Compressor} & & & & & & & \\ 
		& 	$T_{air,in}$  		& $^o$C & M  & &  & & & \\ 
		&	$T_{air,out}$  		& $^o$C & C  & & & & &  \\ 
		&	$\dot{m}_{air}$  	& $kg/s$ & C  & & & & &  \\
		&   $\beta_{comp}$		& - 	& M & & & & &  \\
		\multicolumn{2}{l}{Turbine} & & & & & & &  \\
		&	$T_{air,in}$   		& $^o$C & M  & & & & &  \\ 
		&	$T_{air,out}$  		& $^o$C & M  & & & & &  \\ 
		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & & & & &  \\
		&   $\beta_{exp}$		& - 	& C & & & & &  \\
		\multicolumn{2}{l}{Bypass valve} & & & & & & &  \\
		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & & & & &  \\ 
		\multicolumn{2}{l}{Cylinders} & & & & & & &   \\
		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & & & & &  \\
		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & & & & &  \\
		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & & & & &  \\
		&	$T_{air,in}$  		& $^o$C & M  & & & & &  \\ 
		&	$T_{eg,out}$  		& $^o$C & C  & & & & &  \\ 
		\multicolumn{2}{l}{CAC-LT stage} & & & & & & &  \\
		&	$T_{air,in}$  		& $^o$C & C  & & & & &  \\ 
		&	$T_{air,out}$ 		& $^o$C & M  & & & & &  \\ 
		&	$T_{w,in}$   		& $^o$C & M  & & & & &  \\ 
		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & & & & & &  \\
		\multicolumn{2}{l}{CAC-HT stage} & & & & & & & \\
		&	$T_{air,in}$   		& $^o$C & C  & & & & &  \\ 
		&	$T_{air,out}$   	& $^o$C & C  & & & & &  \\ 
		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & & & & &  \\
		\multicolumn{2}{l}{LOC} & & & & & & &  \\
		&	$T_{lo,in}$   		& $^o$C & C  & & & & &  \\  
		&	$T_{lo,out}$    	& $^o$C & M  & & & & &  \\ 
		&	$T_{w,in}$    		& $^o$C & C  & & & & &  \\ 
		&	$T_{w,out}$   		& $^o$C & C  & & & & &  \\ 
		\multicolumn{2}{l}{JWC} & & & & & & &  \\
		&	$T_{wall}$    		& $^o$C & C  & & & & &  \\ 
		&	$T_{w,in}$    		& $^o$C & M  & & & & &  \\ 
		&	$T_{w,out}$	   		& $^o$C & C  & & & & &  \\ 
		& 	$\dot{m}_{w}$  		& $kg/s$ & C & & & & &  \\
		\hline
	\end{tabular}
	\caption{Auxiliary engines, measured and calculated temperatures and flows at different engine loads}
\end{table}

INCLUDE VALIDATION AGAINST PROJECT GUIDE DATA

\begin{table}
	\small
	\centering
	\begin{tabular}{lcrrrrr}
		\hline
		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
		\hline
		Power output & Mech &  &  &  &  &  \\
		Exhaust gas & Heat &  &  &  &  &  \\
		CAC-LT & Heat &  &  &  &  &  \\
		CAC-HT & Heat &  &  &  &  &  \\
		JWC & Heat &  &  &  &  &  \\
		LOC & Heat &  &  &  &  &  \\
		\hline
	\end{tabular}
	\caption{Main engines, energy flows at different engine loads}
\end{table}

\begin{table}
	\small
	\centering
	\begin{tabular}{lcrrrrr}
		\hline
		Energy flow & type & 20\% & 40\% & 60\% & 80\% & 100\% \\
		\hline
		Power output & Mech &  &  &  &  &  \\
		Exhaust gas & Heat &  &  &  &  &  \\
		CAC-LT & Heat &  &  &  &  &  \\
		CAC-HT & Heat &  &  &  &  &  \\
		JWC & Heat &  &  &  &  &  \\
		LOC & Heat &  &  &  &  &  \\
		\hline
	\end{tabular}
	\caption{Auxiliary engines, energy flows at different engine loads}
\end{table}

\subsubsection{Cooling systems}

The ship's cooling systems are designed similarly to most ships and divided in two cooling systems of two separate temperature levels: the high temperature (HT) cooling systems (70-90$^o$C) and the low temperature (LT) cooling systems (40-60$^o$C). The cooling systems provide the necessary cooling to the engines, to the steam systems and to all other components on board. A representation of the ship cooling systems is provided in Figure XXX.

All values not directly measured in the cooling systems were calculated based on mass and energy balances.

Figure cooling systems

@@

\subsubsection{HVAC systems}

@@

\subsubsection{Thrusters}

When entering port areas, the ship needs to use thrusters to maneuver and berth. In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant. In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average. From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).

@@

As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.

The heat demand and generation can be summarized according to the following equations:
\begin{eqnarray}
\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \\
\end{eqnarray}

The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}

\begin{figure}
	\centering
	\includegraphics[width=0.999\linewidth]{Figures/Heating_cooling_systems}
	\caption{General view of the ship heating and cooling systems}
	\label{fig:ship:general}
\end{figure}

\subsubsection{Heat balance parameter estimation}

As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters. The parameter estimation problem is hence written as a minimization problem:
\begin{eqnarray}
min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
\end{eqnarray}

where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections. A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
\begin{table}
	\centering
	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
		\hline 
		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
		\hline
		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
		\hline
	\end{tabular}
	\caption{Parameters optimized in the parameter estimation for the heat balance}
	\label{tab:ParameterEstimation} 
\end{table}

\subsubsection{Heat demand}

We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation. As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
\begin{table}
	\centering
	{\tablinesep=2ex\tabcolsep=10pt
		\begin{tabular}{p{2.8cm}l}
			\hline 
			Heat flow name & Equation \\
			\hline
			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
			\hline
	\end{tabular}}
	\caption{Summary of the heat demand contributions and their calculation}
	\label{tab:HeatDemand}
\end{table}

where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}). The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively. The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.

\subsubsection{Heat generation}

The heat recovered in the EGBs is the only contribution to the heat balance that is known with a reasonable certainty. The heat transferred from the exhaust gas to the steam ($\dot{Q}_{EGB}$) is calculated according to equation \ref{eq:egb}:
\begin{equation}
\dot{Q}_{EGB} = \dot{m}_{eg} c_{p,eg} (T_{eg,EGB,in} - T_{eg,EGB,out})
\end{equation}\label{eq:egb}

where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF

It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines. However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing. In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},

\begin{eqnarray}
\dot{Q}_{HTHR} & = & \dot{Q}_{HTHR,ER1} + \dot{Q}_{HTHR,ER2} \label{eqn:HTHR1} \\
& = & \sum_{i=ER1,ER2}{\epsilon_{HTHR} * \dot{m}_{min,HTHR,i} * c_{p,w} * (T_{HT,out,i} - T_{HRHT,i,in})} \label{eqn:HTHR2}
\end{eqnarray}

where the effectiveness of the heat exchanger $\epsilon_{HTHR}$ is considered to be constant and its value is part of the parameter estimation problem (see table \ref{tab:ParameterEstimation}). The HT water outlet temperature for each engine room is calculated based on the thermal balance of the engines, and the HR water at the HTHR inlet ($T_{HRHT,ER1,in}$) is considered as a calibration parameter.
